---
title: "Chapter 4"
author: "Matt Emery"
date: "October 22, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**4E1.** In the model definition below, which line is the likelihood?

$$y_i ∼ Normal(μ, σ)$$
$$μ ∼ Normal(0, 10)$$

$$σ ∼ Uniform(0, 10)$$
$y_i ∼Normal(\mu, \sigma)$ is the likelihood.

**4E2.** In the model definition just above, how many parameters are in the posterior distribution?

There are two paramaters in the posterior distribution: $\mu$ and $\sigma$.

**4E3.** Using the model definition above, write down the appropriate form of Bayes’ theorem that
includes the proper likelihood and priors.

$$p(\mu,\sigma|y) = \frac{N(y|\mu,\sigma)N(\mu|0, 10)Unif(\sigma|0,10)}{\int\int\ N(y|\mu,\sigma)N(\mu|0, 10)Unif(\sigma|0,10)d\mu d\sigma}$$

**4E4.** In the model definition below, which line is the linear model?
$$y_i ∼ Normal(μ, σ)$$
$$μ_i = α + βx_i$$
$$ α ∼ Normal(0, 10)$$
$$β ∼ Normal(0, 1)$$
$$σ ∼ Uniform(0, 10)$$
Answer: $μ_i = α + βx_i$

**4E5.** In the model definition just above, how many parameters are in the posterior distribution?

Three: $\alpha$, $\beta$ and $\sigma$

## Medium.

**4M1.** For the model definition below, simulate observed heights from the prior (not the posterior).

$$y_i ∼ Normal(μ, σ)$$
$$μ ∼ Normal(0, 10)$$
$$σ ∼ Uniform(0, 10)$$

```{r}
prior <- rnorm(1000, rnorm(1000, 0, 10), runif(1000, 0, 10))
sample(prior, 100)
```


**4M2.** Translate the model just above into a map formula.

**4M3.** Translate the map model formula below into a mathematical model definition.

```{r}
flist <- alist(
y ~ dnorm( mu , sigma ),
mu <- a + b*x,
a ~ dnorm( 0 , 50 ),
b ~ dunif( 0 , 10 ),
sigma ~ dunif( 0 , 50 )
)
```

**4M4.** A sample of students is measured for height each year for 3 years. After the third year, you want
to fit a linear regression predicting height using year as a predictor. Write down the mathematical
model definition for this regression, using any variable names and priors you choose. Be prepared to
defend your choice of priors.

**4M5.** Now suppose I tell you that the average height in the first year was 120 cm and that every
student got taller each year. Does this information lead you to change your choice of priors? How?

**4M6.** Now suppose I tell you that the variance among heights for students of the same age is never
more than 64cm. How does this lead you to revise your priors?